Topological Equivariant Artist Model for Visualization Library Architecture

### 2022-05-18 paper/main.tex
sent off draft

\usepackage[inline]{showlabels}

@@

\IEEEraisesectionheading{\section{Introduction}\label{sec:intro}}
\note{please prioritize flagging content that needs major revisions: (1) Wrong (2) Missing (3) Horrific  (4) Incoherant}

\IEEEPARstart{T}

\note{Functionl, Types, What is viz library architecture}
\section{Related Work}
Many visualization researchers have discussed structure preservation in a visualization context. In this work, we build on and generalize the notion that the structure of a dataset can be described as a combination of the topological properties of the dataset and the mathematical structure of the dataset fields.

@@

To bridge the need for flexibility and cohesiveness, we build on the proposal of fiber bundles as a general abstraction model for visualization data, as proposed by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}. Fiber bundles are a mathematical structure that can be used to encode the topological properties of visualization data. There is also a methodology for encoding field types in fiber bundles, as described in Spivak's description of tables as fiber bundles\cite{spivakSIMPLICIALDATABASES}. Fiber bundles are discussed in more detail in \autoref{sec:atct:fiber-bundles}, but broadly they provide a mathematical representation of data that can encode topological properties, field types, and data values in a way that is both uniform and continuity and dimension independent way.

@@

\section{IO: Data, Graphic}
\label{sec:atct}
A goal of this work is to develop guidelines for building composable structure preserving functions for translating data into visual representations. Concepts in algebraic topology and category theory provide a language for formally expressing the structure that these visualizations are expected to preserve. Fiber bundles and sheaves, which come from algebraic topolology, provide a uniform generalizable way to describe data and visualization. Category theory is a method to describe how objects are specified, constrained, and composed \cite{wielsManagementEvolvingSpecifications1998}\; therefore in this work we propose categorical constructions of visualization components as specifications for implementable components.

@@

\begin{figure}[h!]
  \label{fig:atct:fb:graphic}
  \includegraphics[width=1\columnwidth]{fb_rho.png}
  \caption{A bundle representation of a section that generates a scatter marker on a 2D screen. The fiber space encodes the potential color and position of each pixel in a prerender space. The marker is a 2d graphic; therefore the base space that parameterizes the graphic area \dbasec\ is a 2d disc. This section \gsection\ describes how to build a square red marker; a section evaluated on a single point $\gbasepoint \in \gbase$ is an infinitesimally small piece of the square, here illustrated as the small piece of the square outlined in green.  }
\end{figure}

In \autoref{fig:atct:fb:graphic}, the section function $\gsection$ encodes the drawing of a red square marker. The renderer subsamples the position fibers  \textit{x} and \textit{y} of the graphic fiber space \gfiber. The preimage of the section at these coordinates returns a set of coordinates ${\gsection^{-1}_{x,y}} = \{\gbasepoint | \gsection(\gbasepoint)\restriction_{x,y} = \(x,y\)\}$ at the points $\gbasepoint \in \gbase$ in the graphic base topological space. The section is then evaluated at these points $\gsection\restriction_{\gsection^{-1}_{x,y}}$ to return a set of points in the fiber. One of these points, enlarged for visibility, is illustrated as the small square swatch inside the larger marker in \autoref{fig:atct:fb}. The renderer is handed some version of $\gsection$ to produce the final graphic.

@@

\subsection{Sheaves}
\label{sec:atct:sheaves}
Sheaves on bundles are an "algebraic data structure", as described by Ghrist \cite{ghristElementaryAppliedTopology2014}, for expressing the bookkeeping involved in keeping track of sections of data over opensets. Sheaves are a way of expressing that subsets, distributed data, and streaming data are local sections that can be glued together. Formally, pre-sheaves are functors, meaning they are functions from objects of one category to objects of another category \cite{WhatFunctorDefinitions}. Here the sheaf 
\begin{equation}
  \sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
\end{equation}
denotes a sheaf from opensets on the base space $\dbase$ to sets of sections on the bundle $\dtotal$. As a contravariant functor it preserves morphism s

@@

\begin{figure}[h!]
  \label{fig:atct:phi}
  \includegraphics[width=\columnwidth]{phi.png}
  \caption{Values in a data set can be transformed in three ways: $\dfunch$-values can change position, .e.g transposed;  $\dfunct$-values can change, e.g. doubled; $\dfunc$ - values can change position and value}
\end{figure}
\autoref{fig:atct:phi} provides an example of a base space change \dfunch, a fiber space change \dfunct, and a composition of the two \dfunc\ applied to each data point $x_{\dbasepoint} \in \texttt{data}$. As defined in \autoref{eq:atct:morphism:verify_base}, 
since transposition is a $(\dfunch, \dfuncpull)$, every data value $x_{\dbasepoint}$ must be equal to the value at the corresponding location in the transposed matrix $x_{\dbasepoint} = x_{\dbasepoint^{\prime}}, \dfunch(\dbasepoint^{\prime}) = \dbasepoint$ for all $x_{\dbasepoint^{\prime}} \in texttt{data.T}$ . In \autoref{fig:atct:phi}, this is demonstrated in the columns becoming rows such that there are no new values and each value retains the same neighboring values. As defined in \autoref{eq:atct:morphism:fiber} scaling is a transform $\dfunct$ that changes the value of each point but not its location. In \autoref{fig:atct:phi}, this means that $x_{\dbasepoint^{\prime}}= 2*x_{\dbasepoint}, \dbasepoint = \dbasepoint_{\prime}$ for all indices in \textt{data}, \textt{data.T}. As defined in \autoref{eq:atct:morphism:all} and illustrated in \autoref{fig:atct:phi}, these transformations can combine to yield a matrix that is both transposed and scaled such that $x_{\dbasepoint^{\prime}} = 2*x_{\dbasepoint}$ for all  
$\dfunch(\dbasepoint^{\prime}) = \dbasepoint$

\subsection{Maps Between Spaces}
\label{sec:atct:xi}
The data sections defined in \autoref{eq:atct:fb} are sections of sheaves, introduced in \autoref{sec:atct:sheaves}, over open sets $\openset \subseteq \dbase$ in a base space \dbase\ that encodes the topology of the data topology. The graphic sections, defined in \autoref{eq:atct:fb_graphic_section}, are sections of sheaves over open sets $\opensetg \subseteq \gbase$ in a base space \gbase\ that encodes the topology of the graphic. We propose that there exists a continuous \textcolor{functor}{map between topological spaces} \vindexc 
\begin{equation}
  \label{eq:atct:xi}
  \vindexc: \opensetgc \rightarrow \opensetc 
\end{equation}
such that every $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$.

@@

We propose that visualization libraries implement morphisms from a data sheaf $\sheafc_{\dtotal, \dbase}$ to a graphic sheaf $\sheafc_{\gtotal, \gbase}$ 
\begin{equation}
  \label{eq:atct:sheaves:homset}
  \begin{tikzcd}
    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
    \arrow[dd, "\textcolor{set}{Hom}_{\sheafc_{\dbasec}}"', color=homset] 
    \arrow[rrdd, "\textcolor{set}{Hom}_{\sheafc_{\dbasec},\sheafc_{\gbasec}}", color=homset] 
    \arrow[rr, "\vindexpullc", color=functor] &  &
    \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} 
    \arrow[dd, "\textcolor{set}{Hom}_{\sheafc_{\gbasec}}", color=homset] \\
     & & \\
    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} &  & 
    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
    \arrow[ll, "\vindexpushc"', color=functor]                  
    \end{tikzcd}
\end{equation}
and the hom set is the set of all morphisms from one sheaf to the other. The functors $\vindexpullc, \vindexpushc$ are adjoint, meaning that there is a functorial isomorphism \cite{harder2008lectures} such that 
\begin{equation}
  \label{eq:atct:sheaves:homset:functors}
\begin{split}
  & \textcolor{set}{Hom}_{\sheafc_{\gbasec}}(\vindexpullc\sheafc_{\dtotalc, \dbasec},\sheafc_{\gtotalc,\gbasec})\\
  \simeq & \textcolor{set}{Hom}_{\sheafc_{\dbasec}}(\sheafc_{\dtotalc, \dbasec},\vindexpushc\sheafc_{\gtotalc,\gbasec}) \\
  \simeq & \textcolor{set}{Hom}_{\sheafc_{\dbasec},\sheafc_{\gbasec}}(\sheafc_{\dtotalc, \dbasec},\sheafc_{\gtotalc,\gbasec}) \\
\end{split} 
\end{equation}
which means that the diagram of sheaf morphisms \autoref{eq:atct:sheaves:homset} is commutative; therefore the functors $\vindexpullc$ and $\vindexpushc$ can be used to adapt morphisms written over one space to another space. 
This provides us flexibility over which space to construct data to graphic transforms. For example, specifications such as svg\cite{quintScalable2003} and Vega\cite{satyanarayanDeclarativeInteractionDesign2014}, typically define their transformations in data space and return specifications for the render to evaluate. An artist may also want to specify the transformations in a proxy of graphic space, for example for operations like dynamic resampling \note{citation:root maybe}.

\section{Artist: Data to Graphic}
\label{sec:artist}
In this work we propose that visualization libraries are implementing a subset of the functions in the hom sets in \autoref{eq:atct:sheaves:homset} and we call these subset of functions the artist:
\begin{align}
  \label{eq:artist:hom_transport}
  \vartistc:& \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}\textcolor{artist}{\rightarrow} \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \in \textcolor{set}{Hom}_{\sheafc_{\dbase}, \sheafc_{\gbase}}\\
  \vartistc^{\dbasec}:&\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}  \textcolor{artist}{\rightarrow} \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} \in \textcolor{set}{Hom}{\sheafc_{\dbase}}\\
  \vartistc^{\gbasec}:& \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} \textcolor{artist}{\rightarrow} \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \in \textcolor{set}{Hom}_{\sheafc_{\gbase}}
\end{align}
 Because the artists can be constructed as morphisms of sheave over the same base spaces, as listed in \autoref{eq:atct:sheaves:homset:functors}, they are natural transformations
\begin{align}
  \label{eq:artist:natural_transform}
  \vartistc^{\dbasec}:& \sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow} \vindexpushc \sheafc_{\gbasec, \gtotalc}\\
  \vartistc^{\gbasec}:&\vindexpullc\sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow}\sheafc_{\gbasec, \gtotalc}
\end{align}
which means that they are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}. As illustrated in \autoref{eq:atct:sheaves:homset}, the sheaf functors
\begin{equation}
    \begin{tikzcd}
      \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} &  & \opensetc \arrow[ll, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] \arrow[rr, "{\vindexpushc\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] &  & \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} 
      \end{tikzcd}
\end{equation}
take as input the same object \opensetc\ and return sets of data and graphic sections that are objects in \setb. As a map between these sheaf functors, the artist has to preserve the $\iota, \iota^*$ morphisms of the presheaf functor, described in \autoref{eq:atct:presheaf}, such that the following diagram commutes

\begin{equation}
  \label{eq:artist:natural_transform:inclusions}
  \begin{tikzcd}
    \opensetc_1 & \cgamma{\opensetc_{1}}{\dtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*"', color=set] \arrow[rr, "\vartistc_{\opensetc_1}", color=artist] &  & \cgamma{\opensetc_{1}}{\vindexpushc\gtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*", color=set] \\
      &  &  &  \\
    \opensetc_{2} \arrow[uu, "\iota", hook, color=base] & \cgamma{\opensetc_{2}}{\dtotalc\restruction_{\opensetc_2}} \arrow[rr, "\vartistc_{\opensetc_2}", color=artist]                        &  & \cgamma{\opensetc_{2}}{\vindexpushc\gtotalc\restruction_{\opensetc_2}}                      
    \end{tikzcd}
\end{equation}
 The diagram in \autoref{eq:artist:natural_transform} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace. By construction, the artist is expected to preserve the sheaf axioms of locality and gluing (\autoref{def:atct:sheaf:gluing}, \autoref{def:atct:sheaf:locality}). This bookkeeping is necessary for any visualization technique that selectively acts on different pieces of a data set; for example streaming visualizations \cite{krstajicVisualizationStreamingData2013} and panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}

The output of an artist \vartist\ is a restricted subset of graphic sections
\begin{equation}
  \begin{split}
  \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} &\coloneqq\\ 
  &\{\gsectionc\vert\;\exists\;\dsectionc \in \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}\;s.t.\; 
  \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\opensetgc) = \opensetc \} 
  \end{split}
\end{equation} 
that are, by definition, only reachable through a structure preserving artist, which we describe in \autoref{sec:artist:equivariance}. We define this subset because the space of all sections $\cgamma{\opensetg}{\gtotal\restriction_{\openset}}$ includes sections that may not be structure preserving. For example, a section may go from every point in the graphic space to the same single point in the graphic fiber $\gsection(\gbasepoint_i) = d\; \forall \gbasepoint \in \gbase$ such that the visual output is a single inked pixel on a screen.

@@

We formalize this structure preservation as equivariance, which is that for every morphism on the data $(\dfunch_{\dtotal}, \dfunct_{\dtotal})$ there is an equivalent morphism on the graphic  $(\dfunch_{\gtotal}, \dfunct_{\gtotal})$ The artist is an equivariant map if the diagram commutes for all points $\gbasepointc^{\prime}\in \gbasec^{\prime}$
\label{sec:artist:equivariance}
\begin{equation}
  \autoref{eq:artist:equivariance}
  \begin{tikzcd}[ampersand replacement=\&, column sep=small]
  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
  \arrow[rrr, "\vartistc", color=artist] 
  \arrow[d, "\dfuncpullc_{\dtotalc}"', color=action] 
  \& \& \& 
  \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
  \arrow[d, "\dfuncpullc_{\gtotalc}", dotted] \\
  \cgamma{\opensetc^{\prime}}{\dfuncpullc_{\dtotalc}\dtotalc\restriction_{\opensetc^{\prime}}} 
  \arrow[dd, "\dfunctc_{\dtotalc}"', color=action] \& 
  \opensetc 
   \& 
  \opensetgc 
  \arrow[l, "\vindexc"', color=functor] 
  \& 
  \imartist{\opensetgc^{\prime}}{\dfuncpullc_{\gtotalc}\gtotalc\restriction_{\opensetgc^{\prime}}} 
  \arrow[dd, "\dfunctc_{\gtotalc}", dotted, color=action] \\
  \& 
  \opensetc^{\prime} 
  \arrow[u, "\dfunchc_{\dtotalc}", color=action] 
  \& 
  \opensetgc^{\prime} 
  \arrow[l, "\vindexc"', color=functor] 
  \arrow[u, "\dfunchc_{\gtotalc}"', dotted, color=action] 
  \& \\
  \cgamma{\opensetc^{\prime}}{\dtotalc^{\prime}\restriction_{\opensetc^{\prime}}} 
  \arrow[rrr, "\vartistc^{\prime}", color=artist]  
  \& \& \& 
  \imartist{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}}
  \end{tikzcd}
\end{equation}
such that starting at an arbitrary data point $\dsectionc(\dbasepointc)$ and transforming it into a different data point and then into a graphic 
\begin{equation*}
  \vartistc^{\prime}(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vartistc(\dsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))
\end{equation*}
is equivalent to transforming the original data point into a graphic and then transforming the graphic into another graphic. The function $\dfunch_{\gtotal}$ induces a change in graphic generating function that matches the change in data.

>>>>

%\usepackage[inline]{showlabels}

@@

\IEEEraisesectionheading{\section{Introduction}\label{sec:intro}}
\note{please prioritize flagging content that needs major revisions: (1) Wrong (2) Missing (3) Horrific  (4) Incoherant (5) Movable to appendix - TCVG page limit is 12 pages content + supplemental}

\IEEEPARstart{V}isualization software libraries consist of a set of functions for translating data into visual representations. These functions are expected to preserve the structure of the input data, however that structure is specified. Over time, library developers may need to develop new components to adapt to changing data needs, especially the growth of complex structured and distributed data. We propose that the complex structure of data can be described using concepts from algebraic topology and category theory, and that category theory can be used to express the constraints these components must satisfy in order to produce structure preserving data to visual transformations. This abstraction of the data structure and specificaton of the components can then be used to guide the developement of functional\footnote{functional programming paradigm} components. Since these components are functional, they are expected to be more composable and modular than existing components, have fewer side effects and therefore be easier to incorporate alignside existing library elements, and be verifiable and therefore easier to test \cite{huHowFunctionalProgramming2015,hughesWhyFunctionalProgramming1989}. In this paper, we introduce a mathematical framework for describing, constructing, and testing visualization library components and provide a toy example illustrating these concepts.

\section{Related Work}
Visualization library components are required to preserve structure; in a visualization context this has widely come to be defined as a combination of the topological properties of the dataset and the mathematical structure of the dataset fields.

@@

\note{maybe move the following to \autoref{sec:atct:io} and consolidate?}
To bridge the need for flexibility and cohesiveness, we build on the proposal of fiber bundles as a general abstraction model for visualization data, as proposed by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}. Fiber bundles are a mathematical structure that can be used to encode the topological properties of visualization data. There is also a methodology for encoding field types in fiber bundles, as described in Spivak's description of tables as fiber bundles\cite{spivakSIMPLICIALDATABASES}. Fiber bundles are discussed in more detail in \autoref{sec:atct:fiber-bundles}, but broadly they provide a mathematical representation of data that can encode topological properties, field types, and data values in a way that is both uniform and continuity and dimension independent way.

@@

\section{Bundles and Sheaves}
\label{sec:atct}
A goal of this work is to develop guidelines for building composable structure preserving functions for translating data into visual representations. Concepts in algebraic topology and category theory provide a language for formally expressing the structure that these visualizations are expected to preserve. Fiber bundles and sheaves, which come from algebraic topolology, provide a uniform generalizable way to describe data and visualization. Category theory is a method to describe how objects are specified, constrained, and composed \cite{wielsManagementEvolvingSpecifications1998}\; therefore in this work we propose categorical constructions of visualization components as specifications for implementable components. A brief introduction to category theory for visualization practitioners is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.

@@

\begin{figure}[h!]
  \label{fig:atct:fb:graphic}
  \includegraphics[width=1\columnwidth]{fb_rho.png}
  \caption{A bundle representation of a section that generates a scatter marker on a 2D screen. The fiber space encodes the potential color and position of each pixel in a prerender space. The marker is a 2d graphic; therefore the base space that parameterizes the graphic area \dbasec\ is a 2d disc. This section \gsection\ describes how to build a square red marker; a section evaluated on a single point $\gbasepoint \in \gbase$ is an infinitesimally small piece of the square, here illustrated as the small piece of the square outlined in green.}
\end{figure}

In \autoref{fig:atct:fb:graphic}, the section function $\gsection$ encodes the drawing of a red square marker. The renderer subsamples the position fibers  \textit{x} and \textit{y} of the graphic fiber space \gfiber. The preimage of the section at these coordinates returns a set of coordinates \({\gsection^{-1}_{x,y}} = \{\gbasepoint | \gsection(\gbasepoint)\restriction_{x,y} = \(x,y\)\}\) at the points \(\gbasepoint \in \gbase\) in the graphic base topological space. The section is then evaluated at these points \(\gsection\restriction_{\gsection^{-1}_{x,y}}\) to return a set of points in the fiber. One of these points, enlarged for visibility, is illustrated as the small square swatch inside the larger marker in \autoref{fig:atct:fb}. The renderer is handed some version of $\gsection$ to produce the final graphic.

@@

\subsection{Sheaves}
\label{sec:atct:sheaves}
Sheaves on bundles are an "algebraic data structure", as described by Ghrist \cite{ghristElementaryAppliedTopology2014}, for expressing the bookkeeping involved in keeping track of sections of data over opensets. Sheaves are a way of expressing that subsets, distributed data, and streaming data are local sections that can be glued together. Formally, pre-sheaves are functors, meaning they are functions from objects of one category to objects of another category \cite{WhatFunctorDefinitions}. Here the sheaf 
\begin{equation}
  \label{eq:atct:sheaf:functor}
  \sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
\end{equation}
denotes a sheaf from opensets on the base space $\dbase$ to sets of sections on the bundle $\dtotal$. As a contravariant functor it preserves morphism s

@@

\begin{figure}[h!]
  \label{fig:atct:phi}
  \includegraphics[width=\columnwidth]{phi.png}
  \caption{Values in a data set can be transformed in three ways: $\dfunch$-values can change position, .e.g transposed;  $\dfunct$-values can change, e.g. doubled; $\dfunc$ - values can change position and value}
\end{figure}
\autoref{fig:atct:phi} provides an example of a base space change \dfunch, a fiber space change \dfunct, and a composition of the two \dfunc\ applied to each data point $x_{\dbasepoint} \in \texttt{data}$. As defined in \autoref{eq:atct:morphism:verify_base}, 
since transposition is a $(\dfunch, \dfuncpull)$, every data value $x_{\dbasepoint}$ must be equal to the value at the corresponding location in the transposed matrix $x_{\dbasepoint} = x_{\dbasepoint^{\prime}}, \dfunch(\dbasepoint^{\prime}) = \dbasepoint$ for all $x_{\dbasepoint^{\prime}} \in texttt{data.T}$ . In \autoref{fig:atct:phi}, this is demonstrated in the columns becoming rows such that there are no new values and each value retains the same neighboring values. As defined in \autoref{eq:atct:morphism:fiber} scaling is a transform $\dfunct$ that changes the value of each point but not its location. In \autoref{fig:atct:phi}, this means that $x_{\dbasepoint^{\prime}}= 2*x_{\dbasepoint}, \dbasepoint = \dbasepoint_{\prime}$ for all indices in \texttt{data}, \texttt{data.T}. As defined in \autoref{eq:atct:morphism:all} and illustrated in \autoref{fig:atct:phi}, these transformations can combine to yield a matrix that is both transposed and scaled such that $x_{\dbasepoint^{\prime}} = 2*x_{\dbasepoint}$ for all  
$\dfunch(\dbasepoint^{\prime}) = \dbasepoint$

\subsection{Image Functors}
\label{sec:atct:xi}
The data sections defined in \autoref{eq:atct:fb} are sections of sheaves, introduced in \autoref{sec:atct:sheaves}, over open sets $\openset \subseteq \dbase$ in a base space \dbase\ that encodes the topology of the data topology. The graphic sections, defined in \autoref{eq:atct:fb_graphic_section}, are sections of sheaves over open sets $\opensetg \subseteq \gbase$ in a base space \gbase\ that encodes the topology of the graphic. We propose that there exists a continuous \textcolor{functor}{map between topological spaces} \vindexc 
\begin{equation}
  \label{eq:atct:xi}
  \vindexc: \opensetgc \rightarrow \opensetc 
\end{equation}
such that every $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$.

@@

We propose that visualization libraries implement morphisms from a data sheaf $\sheafc_{\dtotal, \dbase}$ to a graphic sheaf $\sheafc_{\gtotal, \gbase}$ 
\begin{equation}
  \label{eq:atct:sheaves:homset}
  \begin{tikzcd}
    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
    \arrow[dd, "\textcolor{set}{Hom}_{\sheafc_{\dbasec}}"', color=homset] 
    \arrow[rrdd, "\textcolor{set}{Hom}_{\sheafc_{\dbasec},\sheafc_{\gbasec}}", color=homset] 
    \arrow[rr, "\vindexpullc", color=functor] &  &
    \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} 
    \arrow[dd, "\textcolor{set}{Hom}_{\sheafc_{\gbasec}}", color=homset] \\
     & & \\
    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} &  & 
    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
    \arrow[ll, "\vindexpushc"', color=functor]                  
    \end{tikzcd}
\end{equation}
and the hom set is the set of all morphisms from one sheaf to the other. The functors $\vindexpullc, \vindexpushc$ are adjoint, meaning that there is a functorial isomorphism \cite{harder2008lectures} such that 
\begin{equation}
  \label{eq:atct:sheaves:homset:functors}
\begin{split}
  & \textcolor{set}{Hom}_{\sheafc_{\gbasec}}(\vindexpullc\sheafc_{\dtotalc, \dbasec},\sheafc_{\gtotalc,\gbasec})\\
  \simeq & \textcolor{set}{Hom}_{\sheafc_{\dbasec}}(\sheafc_{\dtotalc, \dbasec},\vindexpushc\sheafc_{\gtotalc,\gbasec}) \\
  \simeq & \textcolor{set}{Hom}_{\sheafc_{\dbasec},\sheafc_{\gbasec}}(\sheafc_{\dtotalc, \dbasec},\sheafc_{\gtotalc,\gbasec}) \\
\end{split} 
\end{equation}
which means that the diagram of sheaf morphisms \autoref{eq:atct:sheaves:homset} is commutative; therefore the functors $\vindexpullc$ and $\vindexpushc$ can be used to adapt morphisms written over one space to another space. 
This provides us flexibility over which space to construct data to graphic transforms. For example, specifications such as svg\cite{quintScalable2003} and Vega\cite{satyanarayanDeclarativeInteractionDesign2014}, typically define their transformations in data space and return specifications for the render to evaluate. An artist may also want to specify the transformations in a proxy of graphic space, for example to generate on-demand dynamically resampled continuous lines.

\section{Artist: Data to Graphic}
\label{sec:artist}
In this work we propose that visualization libraries are implementing a subset of the functions in the hom sets in \autoref{eq:atct:sheaves:homset}. We call these subset of functions the artist:
\begin{align}
  \label{eq:artist:hom_transport}
  \vartistc:& \cgamma{\dbasec}{\dtotalc}\textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\gtotalc} \in \textcolor{set}{Hom}_{\sheafc_{\dbase}, \sheafc_{\gbase}}\\
  \vartistc^{\dbasec}:&\cgamma{\dbasec}{\dtotalc}  \textcolor{artist}{\rightarrow} \cgamma{\dbase}{\vindexpushc\gtotalc} \in \textcolor{set}{Hom}{\sheafc_{\dbase}}\\
  \vartistc^{\gbasec}:& \cgamma{\gbasec}{\vindexpullc\dtotalc} \textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\gtotalc} \in \textcolor{set}{Hom}_{\sheafc_{\gbase}}
\end{align}
 Because the artists can be constructed as morphisms of sheave over the same base spaces, as listed in \autoref{eq:atct:sheaves:homset:functors}, they are natural transformations
\begin{align}
  \label{eq:artist:natural_transform}
  \vartistc^{\dbasec}:& \sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow} \vindexpushc \sheafc_{\gbasec, \gtotalc}\\
  \vartistc^{\gbasec}:&\vindexpullc\sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow}\sheafc_{\gbasec, \gtotalc}
\end{align}
which means that they are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}. As illustrated in \autoref{eq:atct:sheaves:homset}, the sheaf functors
\begin{equation}
  \label{eq:artist:sheaf:base}
    \begin{tikzcd}
      \cgamma{\dbasec}{\dtotalc} &  & \dbasec \arrow[ll, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] \arrow[rr, "{\vindexpushc\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] &  & \cgamma{\dbase}{\vindexpushc\gtotalc} 
      \end{tikzcd}
\end{equation}
take as input the same object \opensetc\ and return sets of data and graphic sections that are objects in \setb. As a map between these sheaf functors, the artist has to preserve the $\iota, \iota^*$ morphisms of the presheaf functor, described in \autoref{eq:atct:presheaf}, such that the following diagram commutes

\begin{equation}
  \label{eq:artist:natural_transform:inclusions}
  \begin{tikzcd}
    \dbasec_1 & \cgamma{\dbasec_1}{\dtotalc} 
    \arrow[dd, "\iota^*"', color=set] 
    \arrow[rr, "\vartistc_{\dbase_1}", color=artist] &  & 
    \cgamma{\dbasec_{1}}{\vindexpushc\gtotalc} 
    \arrow[dd, "\iota^*", color=set] \\
      &  &  &  \\
    \dbasec_{2} \arrow[uu, "\iota", hook, color=base] & 
    \cgamma{\dbasec_{2}}{\dtotalc} 
    \arrow[rr, "\vartistc_{\dbasec_2}", color=artist] &  & 
    \cgamma{\dbasec_{2}}{\vindexpushc\gtotalc}                      
    \end{tikzcd}
\end{equation}
 The diagram in \autoref{eq:artist:natural_transform} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace. By construction, the artist is expected to preserve the sheaf axioms of locality and gluing (\autoref{def:atct:sheaf:gluing}, \autoref{def:atct:sheaf:locality}). This bookkeeping is necessary for any visualization technique that selectively acts on different pieces of a data set; for example streaming visualizations \cite{krstajicVisualizationStreamingData2013} and panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}

The output of an artist \vartist\ is a restricted subset of graphic sections
\begin{equation}
  \label{eq:artist:output}
  \imartist{\gbasec}{\gtotalc} \coloneqq\\ 
  \{\gsectionc \mid\;\exists\;\dsectionc \in \cgamma{\dbasec}{\dtotalc}\;s.t.\; 
  \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\gbasec) = \dbasec \}
\end{equation} 
that are, by definition, only reachable through a structure preserving artist, which we describe in \autoref{sec:artist:equivariance}. We define this subset because the space of all sections $\cgamma{\opensetg}{\gtotal\restriction_{\openset}}$ includes sections that may not be structure preserving. For example, a section may go from every point in the graphic space to the same single point in the graphic fiber $\gsection(\gbasepoint_i) = d\; \forall \gbasepoint \in \gbase$ such that the visual output is a single inked pixel on a screen.

@@

We formalize this structure preservation as equivariance, which is that for every morphism on the data $(\dfunch_{\dtotal}, \dfunct_{\dtotal})$ there is an equivalent morphism on the graphic  $(\dfunch_{\gtotal}, \dfunct_{\gtotal})$ The artist is an equivariant map if the diagram commutes for all points $\gbasepointc^{\prime}\in \gbasec^{\prime}$
\label{sec:artist:equivariance}
\begin{equation}
  \autoref{eq:artist:equivariance}
  \begin{tikzcd}[ampersand replacement=\&, column sep=small]
  \cgamma{\dbasec}{\dtotalc} 
  \arrow[rrr, "\vartistc", color=artist] 
  \arrow[d, "\dfuncpullc_{\dtotalc}"', color=action] 
  \& \& \& 
  \imartist{\gbasec}{\gtotalc} 
  \arrow[d, "\dfuncpullc_{\gtotalc}", dotted] \\
  \cgamma{\dbasec^{\prime}}{\dfuncpullc_{\dtotalc}\dtotalc} 
  \arrow[dd, "\dfunctc_{\dtotalc}"', color=action] \& 
  \opensetc 
   \& 
  \opensetgc 
  \arrow[l, "\vindexc"', color=functor] 
  \& 
  \imartist{\gbase^{\prime}}{\dfuncpullc_{\gtotalc}\gtotalc} 
  \arrow[dd, "\dfunctc_{\gtotalc}", dotted, color=action] \\
  \& 
  \opensetc^{\prime} 
  \arrow[u, "\dfunchc_{\dtotalc}", color=action] 
  \& 
  \opensetgc^{\prime} 
  \arrow[l, "\vindexc"', color=functor] 
  \arrow[u, "\dfunchc_{\gtotalc}"', dotted, color=action] 
  \& \\
  \cgamma{\dbasec^{\prime}}{\dtotalc^{\prime}} 
  \arrow[rrr, "\vartistc^{\prime}", color=artist]  
  \& \& \& 
  \imartist{\gbasec^{\prime}}{\gtotalc^{\prime}}
  \end{tikzcd}
\end{equation}
such that starting at an arbitrary data point $\dsectionc(\dbasepointc)$ and transforming it into a different data point and then into a graphic 
\begin{equation*}
  \vartistc^{\prime}(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vartistc(\dsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))
\end{equation*}
is equivalent to transforming the original data point into a graphic and then transforming the graphic into another graphic. The function $\dfunch_{\gtotal}$ induces a change in graphic generating function that matches the change in data.
